Learning From Massive Noisy Labeled Data for Image Classification

### 2015-04-09 paper.tex
Fix the bibs

Deep learning with a massive amount of supervised training data has recently shown very impressive improvement on multiple image recognition challenges including image classification~\cite{krizhevsky2012imagenet}, attribute learning~\cite{zhang2013panda}, and scene classification~\cite{farabet2013learning}. While state-of-the-art results have been continuously reported~\cite{zeiler2013visualizing,simonyan2014very,szegedy2014going}, all these methods require reliable annotations from millions of images~\cite{deng2009imagenet} which are often expensive and time-consuming to obtain~\cite{deng2009imagenet}, preventing deep models from being quickly trained on new image recognition problems. Thus it is necessary to develop new efficient labeling and training frameworks for deep learning.

@@

Although annotating all the data is costly, it is often easy to obtain a small amount of clean labels. Based on the observation of transferability of deep neural networks, people initialize parameters with a model pretrained on a larger yet related dataset~\cite{krizhevsky2012imagenet}, and then finetune on the smaller dataset of specific tasks~\cite{oquab2013learning,azizpour2014generic,donahue2013decaf}. Such methods may better avoid overfitting and utilize the relationships between the two datasets. However, we find that training a CNN from scratch with limited clean labels and massive noisy labels is better than finetuning it only on clean labels. Other approaches address the problem as semi-supervised learning where noisy labels are discarded~\cite{zhu2002learning}. These algorithms usually suffer from model complexity thus cannot be applied on large-scale datasets. Therefore, it is inevitable to develop a better way of using huge amount of noisy labeled data.

@@

To cope with this challenge, we extend CNNs with a novel probabilistic model, which infers the true labels and uses them to supervise the training of the network. Our work is inspired by~\cite{sukhbaatar2014learning}, which modified a CNN by inserting a linear layer on top of the softmax layer to map clean labels to noisy labels. However,~\cite{sukhbaatar2014learning} assumed noisy labels were conditionally independent with input images given clean labels. By examining our collected dataset, we find that this assumption is too strong to fit real-world data well. For example, in Figure~\ref{fig:noise_cluster}, all the images should belong to ``Hoodie''. The top five are correct while the bottom five are either mislabeled as ``Windbreaker'' or ``Jacket''. It shows that images tend to be mislabeled may share similar visual patterns. This is because different sellers have their own bias on different categories, thus they may provide wrong keywords for many similar clothes. Based on these observations, we introduce two types of label noise:
\begin{itemize}
\item \textbf{Confusing noise} makes the noisy label reasonably wrong. It usually occurs when the image content is confusing (\eg, the samples with ``?'' in Figure~\ref{fig:overview}).
\item \textbf{Pure random noise} makes the noisy label totally wrong. It is often caused by either the mismatch between an image and its surrounding text, or false conversion from the text to label (\eg, the samples with ``$\times$'' in Figure~\ref{fig:overview}).
\end{itemize}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=1.0\linewidth]{figure/noise_cluster.pdf}
\end{center}
\caption{Images tend to be mislabeled often share similar visual patterns.}
\label{fig:noise_cluster}
\end{figure}
% section introduction (end)

\section{Related Work} % (fold)
\label{sec:related_work}
For most of the related works including the effect of label noises, taxonomy of label noises, robust algorithms and noise cleaning algorithms for learning with noisy data, we refer to~\cite{frenay2013classification} for a comprehensive review.

{\bf Direct learning with noisy labels:} Many works have shown that label noise can adversely impact the classification accuracy of induced classifiers~\cite{zhu2004class}. To better handle label noise, some approaches rely on training classifiers with label noise-robust algorithms~\cite{beigman2009learning,manwani2013noise}. However, Bartlett \etal~\cite{bartlett2006convexity} proved that most of the loss functions are not completely robust to label noise. Experiments in~\cite{teng2001comparison} showed that the classifiers inferred by label noise-robust algorithms are still affected by label noise. These methods seem to be adequate only when label noise can be safely managed by overfitting avoidance~\cite{frenay2013classification}. On the other hand, some label noise cleansing methods were proposed to remove or correct mislabeled instances~\cite{barandela2000decontamination,brodley2011identifying,miranda2009use}, but these approaches were difficult in distinguishing informative hard examples from harmful mislabeled ones. Thus they might remove too many instances and the overcleansing could reduce the performances of classifiers~\cite{matic1992computer}.

@@

{\bf Transfer learning:} The success of CNNs lies in their capability of learning rich and hierarchical image features. However, the model parameters cannot be properly learned when training data is not enough. Researchers proposed to conquer this problem by first initializing CNN parameters with a model pretrained on a larger yet related dataset, and then finetuning it on the smaller dataset of specific task~\cite{krizhevsky2012imagenet, oquab2013learning, azizpour2014generic,donahue2013decaf}. Nevertheless, this transfer learning scheme could be suboptimal when the two tasks are just loosely related. In our case of clothing classification, we find that training a CNN from scratch with limited clean labels and massive noisy labels is better than finetuning it only on the clean labels.

@@

Next we demonstrate the effect of learning to predict the label noise type. We estimate $p(\vec{z}_2=1|\vec{x})+p(\vec{z}_3=1|\vec{x})$ on the validation set and sort the images accordingly in descending order. Then for each image, we check whether its noisy label mismatches its clean label. The rank-precision curve is plotted in Figure~\ref{fig:noise_prediction_rp}. It shows that the intuitive observation --- images tend to be mislabeled often share similar patterns --- is reasonable, and our model is trained to recognize such patterns to help handle noisy labels.

>>>>

Deep learning with a massive amount of supervised training data has recently shown very impressive improvement on multiple image recognition challenges including image classification~\cite{krizhevsky2012imagenet}, attribute learning~\cite{zhang2014panda}, and scene classification~\cite{farabet2013learning}. While state-of-the-art results have been continuously reported~\cite{zeiler2013visualizing,simonyan2014very,szegedy2014going}, all these methods require reliable annotations from millions of images~\cite{deng2009imagenet} which are often expensive and time-consuming to obtain~\cite{deng2009imagenet}, preventing deep models from being quickly trained on new image recognition problems. Thus it is necessary to develop new efficient labeling and training frameworks for deep learning.

@@

Although annotating all the data is costly, it is often easy to obtain a small amount of clean labels. Based on the observation of transferability of deep neural networks, people initialize parameters with a model pretrained on a larger yet related dataset~\cite{krizhevsky2012imagenet}, and then finetune on the smaller dataset of specific tasks~\cite{oquab2014learning,azizpour2014generic,donahue2013decaf}. Such methods may better avoid overfitting and utilize the relationships between the two datasets. However, we find that training a CNN from scratch with limited clean labels and massive noisy labels is better than finetuning it only on clean labels. Other approaches address the problem as semi-supervised learning where noisy labels are discarded~\cite{zhu2002learning}. These algorithms usually suffer from model complexity thus cannot be applied on large-scale datasets. Therefore, it is inevitable to develop a better way of using huge amount of noisy labeled data.

@@

To cope with this challenge, we extend CNNs with a novel probabilistic model, which infers the true labels and uses them to supervise the training of the network. Our work is inspired by~\cite{sukhbaatar2014learning}, which modified a CNN by inserting a linear layer on top of the softmax layer to map clean labels to noisy labels. However,~\cite{sukhbaatar2014learning} assumed noisy labels were conditionally independent with input images given clean labels. By examining our collected dataset, we find that this assumption is too strong to fit real-world data well. For example, in Figure~\ref{fig:noise_cluster}, all the images should belong to ``Hoodie''. The top five are correct while the bottom five are either mislabeled as ``Windbreaker'' or ``Jacket''. This is because different sellers have their own bias on different categories, thus they may provide wrong keywords for similar clothes. We can capture these visual patterns to estimate how likely an image is mislabeled. Based on these observations, we further introduce two types of label noise:
\begin{itemize}
\item \textbf{Confusing noise} makes the noisy label reasonably wrong. It usually occurs when the image content is confusing (\eg, the samples with ``?'' in Figure~\ref{fig:overview}).
\item \textbf{Pure random noise} makes the noisy label totally wrong. It is often caused by either the mismatch between an image and its surrounding text, or false conversion from the text to label (\eg, the samples with ``$\times$'' in Figure~\ref{fig:overview}).
\end{itemize}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=1.0\linewidth]{figure/noise_cluster.pdf}
\end{center}
\caption{Mislabeled images often share similar visual patterns.}
\label{fig:noise_cluster}
\end{figure}
% section introduction (end)

\section{Related Work} % (fold)
\label{sec:related_work}
For most of the related works including the effect of label noises, taxonomy of label noises, robust algorithms and noise cleaning algorithms for learning with noisy data, we refer to~\cite{frenay2014classification} for a comprehensive review.

{\bf Direct learning with noisy labels:} Many works have shown that label noise can adversely impact the classification accuracy of induced classifiers~\cite{zhu2004class}. To better handle label noise, some approaches rely on training classifiers with label noise-robust algorithms~\cite{beigman2009learning,manwani2013noise}. However, Bartlett \etal~\cite{bartlett2006convexity} proved that most of the loss functions are not completely robust to label noise. Experiments in~\cite{teng2001comparison} showed that the classifiers inferred by label noise-robust algorithms are still affected by label noise. These methods seem to be adequate only when label noise can be safely managed by overfitting avoidance~\cite{frenay2014classification}. On the other hand, some label noise cleansing methods were proposed to remove or correct mislabeled instances~\cite{barandela2000decontamination,brodley2011identifying,miranda2009use}, but these approaches were difficult in distinguishing informative hard examples from harmful mislabeled ones. Thus they might remove too many instances and the overcleansing could reduce the performances of classifiers~\cite{matic1992computer}.

@@

{\bf Transfer learning:} The success of CNNs lies in their capability of learning rich and hierarchical image features. However, the model parameters cannot be properly learned when training data is not enough. Researchers proposed to conquer this problem by first initializing CNN parameters with a model pretrained on a larger yet related dataset, and then finetuning it on the smaller dataset of specific task~\cite{krizhevsky2012imagenet, oquab2014learning, azizpour2014generic,donahue2013decaf}. Nevertheless, this transfer learning scheme could be suboptimal when the two tasks are just loosely related. In our case of clothing classification, we find that training a CNN from scratch with limited clean labels and massive noisy labels is better than finetuning it only on the clean labels.

@@

Next we demonstrate the effect of learning to predict the label noise type. We estimate $p(\vec{z}_2=1|\vec{x})+p(\vec{z}_3=1|\vec{x})$ on the validation set and sort the images accordingly in descending order. Then for each image, we check whether its noisy label mismatches its clean label. The rank-precision curve is plotted in Figure~\ref{fig:noise_prediction_rp}. It shows that the intuitive observation --- mislabeled images often share similar patterns --- is reasonable, and our model is trained to recognize such patterns to help handle noisy labels.
